How Short Lived is the Population III Epoch?

### 2010-09-01 ms.tex
Adding references.

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{refs}.  A fraction of these stars enrich the
surrounding intergalactic medium (IGM) when they go supernova, which
can happen in stars $\lsim40~\Ms$ in Type II supernovae (SNe) or in
stars roughly between 140 \Ms~and 260 \Ms~in pair-instability SNe
\citep[PISNe;][]{Heger02}.  The host halo and the neighboring halos
are then enriched with this ejecta.  There exists a critical
metallicity that is $\sim 10^{-6} Z_\odot$ if dust cooling is
efficient \citep{refs} and $\sim 10^{-3.5} Z_\odot$ otherwise
\citep{refs}, where the gas can cool rapidly, lowering its Jeans mass.
An intermediate characteristic mass of $\sim10~\Ms$ can be occur if
the gas cooling is suppressed to the cosmic microwave background (CMB)
temperature \citep{ref}.  The resulting Population II star cluster
will thus have a lower characteristic stellar mass than its metal-free
progenitors \citep{refs}.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = ??? \citep{ref},
and in dwarf spheroidal (dSph) galaxies this floor is a factor of 10
higher \citep{refs}.

The transition from Pop III to Pop II star formation is solely
dependent on the propagation of metals from the SNe remnants into
future sites of star formation.  Their flows are complex because of
the interactions between the SN blastwave, cosmological accretion and
halo mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h}
\sim 10^6~\Ms$), radiation from the massive Pop III drives a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \citep{refs}.  The latter feature aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with an metal bubble size of a few kpc.  The
remaining metals fall back into the host halo.  The blastwave may
induce star formation in nearby halos through the compression of the
gas \citep{refs}; however timescales for metal mixing into the dense
gas are many dynamical times \citep{refs} for shock velocities $\lsim
100~\kms$.

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise11}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = ???$ and use the cosmological parameters from the 7-year WMAP
data \citep{ref}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

We peform two simulations with the same initial conditions.  They
differ only in that one considers primordial cooling, and the other
considers additional cooling in metal-enriched gas.  For primordial
cooling, we use a non-equilbrium chemistry solver with 9 species of
hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
cooling curve that is calculated from CLOUDY \citep{ref} and using the
method of \citet{2008MNRAS.385.1443S}.  It is tabulated based on
density, temperature, electron fraction, and metallicity.

>>>>

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{Abel00, ABN02, Bromm02, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \citep{refs} and $\sim 10^{-3.5}
Z_\odot$ otherwise \citep{Bromm01, 2009ApJ...691..441S}, where the gas
can cool rapidly, lowering its Jeans mass.  An intermediate
characteristic mass of $\sim10~\Ms$ can be occur if the gas cooling is
suppressed to the cosmic microwave background (CMB) temperature
\citep{Larson, Tumlinson04, 2009ApJ...691..441S}.  The resulting
Population II star cluster will thus have a lower characteristic
stellar mass than its metal-free progenitors.  In the local universe,
stars in the Milky Way halo and dwarf spheroidal (dSph) galaxies may
be connected to these first stellar clusters.  In the halo of the
Milky Way, the most metal-poor stars have a metallicity floor of [Z/H]
= ???  \citep{ref}, and in dwarf spheroidal (dSph) galaxies this floor
is a factor of 10 higher \citep{refs}.

The transition from Pop III to Pop II star formation is solely
dependent on the propagation of metals from the SNe remnants into
future sites of star formation.  Their flows are complex because of
the interactions between the SN blastwave, cosmological accretion and
halo mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h}
\sim 10^6~\Ms$), radiation from the massive Pop III drives a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \citep{Kitayama04, Whalen04, Abel07}.  The latter
feature aids in the expansion of the blastwave because it delays the
transition to the Sedov-Taylor and snowplow phases.  In PISNe,
approximately half of the metals stay in the IGM with an metal bubble
size of a few kpc.  The remaining metals fall back into the host halo.
The blastwave may induce star formation in nearby halos through the
compression of the gas \citep{Ferrara98}; however timescales for metal
mixing into the dense gas are many dynamical times \citep{Cen07} for
shock velocities $\lsim 100~\kms$.

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise10}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = ???$ and use the cosmological parameters from the 7-year WMAP
data \citep{ref}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

We peform two simulations with the same initial conditions.  They
differ only in that one considers primordial cooling, and the other
considers additional cooling in metal-enriched gas.  For primordial
cooling, we use a non-equilbrium chemistry solver with 9 species of
hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
cooling curve that is calculated from CLOUDY \citep{CLOUDY} and using
the method of \citet{2008MNRAS.385.1443S}.  It is tabulated based on
density, temperature, electron fraction, and metallicity.
